Behavior Driven Development in Haskell

### 2012-05-25 paper.tex
s/visible/for users/ s/internal/for developers/

\begin{itemize}
\item Introduced {\tt doctest} from Python community to Haskell community
and integrated it with QuickCheck
(Section \ref{sec:Visible-behavior} and Section \ref{sec:doctest})
%% \item Introduced {\tt Hspec} from Ruby community to Haskell community and
%% integrated it with QuickCheck
%% (Section \ref{sec:Internal-behavior} and Section \ref{sec:hspec})
\item Show best current practice of behavior driven development in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
\end{itemize}

@@

We can categorize behaviors into two types: visible behaviors
and internal behaviors.
Visible behaviors are those which are useful for both developers and users.
By contrast, internal behaviors are those which are usefull only for developers.
A good example of internal behaviors
is behavior for a corner case registered in a bug track system.

BDD tools should be treat both of them and provide functionality
to disclose visible behaviors as documentation for users.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

The standard documentation mechanism for Haskell is
Haddock~\cite{haddock}.
Haddock defines markup syntax in Haskell's comment.
Haddock provides code block markup indicated with ({\tt >}).
For example, the {\tt container} packages has the following code block markup:

@@

\begin{itemize}
\item Haskell community does not have common syntax to describe
behaviors in Haddock.
\item Haskell community does not have a BBD tool for visible behaviors.
\item Haskell community does not have a BBD tool for internal behaviors.
\end{itemize}

@@

To specify visible behavior, we propose to add two new markups to Haddock and to use {\tt doctest} integrated with QuickCheck. To describe internal behavior, we propose to use \hspec{} aslo integrated with QuickCheck.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Visible behavior}
\label{sec:Visible-behavior}

\begin{quote}
    @kazu: I think all behavior is user visible, in the way that the user can
    observe it when he uses the function.  Otherwise we would not be interested
    in that behavior at all.  What do you think about using \emph{Behavior as
    user documentation} and \emph{Behavior as code documentation} instead of
    visible/internal behavior?  Or can we come up with something else?
\end{quote}

Useful behavior should be documented in Haddock.
For this purpose, we propose the followings:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Internal behavior}
\label{sec:Internal-behavior}

@@

\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write visible behaviors in Haddock markup:

@@

\noindent We should execute {\tt doctest} to ensure no syntax errors.
Also, we specifies internal behaviors with \hspec{}:

>>>>

\begin{itemize}
\item Introduced {\tt doctest} from Python community to Haskell community
and integrated it with QuickCheck
(Section \ref{sec:for-users} and Section \ref{sec:doctest})
%% \item Introduced {\tt Hspec} from Ruby community to Haskell community and
%% integrated it with QuickCheck
%% (Section \ref{sec:for-developers} and Section \ref{sec:hspec})
\item Show best current practice of behavior driven development in Haskell using {\tt doctest} and \hspec{} (Section \ref{ref:Best-current-practice})
\end{itemize}

@@

We can categorize examples/properties into two types: for users and
and for developers.
Examples/properties for users are those which are useful as user documentation.
By contrast, examples/properties for developers are
those which are not useful as user documentation.
A good example of examples/properties for developers
is a corner example registered in a bug track system.

BDD tools should be treat both of them and provide functionality
to disclose examples/properties for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

The standard documentation mechanism for Haskell is
Haddock~\cite{haddock}.
Haddock defines markup syntax in Haskell's comment.
Haddock provides the code block markup indicated with ({\tt >}).
For example, the {\tt container} packages has the following code block markup:

@@

\begin{itemize}
\item Haskell community does not have common syntax to describe
behaviors in Haddock.
\item Haskell community does not have a BBD tool for 
examples/properties for users.
\end{itemize}

@@

To specify examples/properties for users, we propose to add two new markups to Haddock and to use {\tt doctest} integrated with QuickCheck. To describe examples/properties for developers, we propose to use \hspec{} also integrated with QuickCheck.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Examples/properties for users}
\label{sec:for-users}

Examples/properties for users should be documented with Haddock markup syntax.
For this purpose, we propose the followings:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Examples/properties for developers}
\label{sec:for-developers}

@@

\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write examples/properties for users in Haddock markup:

@@

\noindent We should execute {\tt doctest} to ensure no syntax errors.
Also, we specifies examples/properties for developers with \hspec{}:
